\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-07 arch.tex
Reintroduced reference to IDEA:   - this time in europa reactor section   - I need to read in depth AAAI fall symposium 2006 as they may share a lot of our ideas presented here

The T-REX architecture was designed with the objective to bring 
planning closer to the low-level control of the vehicle while
maintaining its reactivity. Indeed, in most of the architecture where
task planning is embedded, this task remains at a failry high level of
abstraction as the complexity of this task would otherwise impede the
system reactiveness. As a result the planning problem managed {\em in
  situ} remains fairly detached from low-level control and most of the 
reactive adaptation which in contrast is managed by specialized 
component that relies on different representations for modeling 
the local behavior they control or even do not have any explicit 
formal model of this behavior. The two angles our architecture attck
this issue are the following:
\begin{itemize}
\item Distribute the global planning problem into multiple decision
  loops, called {\em reactors}. Each reactor have its own scope both 
  functionally (one for example focusing on the conversion of high 
  level waypoints into low-level commands) and temporally as 
  each reactor specifies to the agent their planning look-ahead and
  their expected latency before being able to produce plan.
\item Enforce a stronger interaction between the planning process 
  and the integration of the world evolution. The objective being to
  make sure that the plannign task is done while continuously informed 
  on new world state change. By doing so, we expect to allow the planner
  tohave better informed decision despite the fact that its latency is
  larger that the world change rate.
\end{itemize}

For the first point T-REX provides a formal framework described in
details in \cite{py10} and is not the focus of this paper. The
second have deep implication on how on should integrate a planner in
this framework in general and was applied to the europa framework
specifically. We'll develop later how this requirement implies to see
in-situ planning as two parallels processes that are tightly linked
through the same plan structure. But before that we need to introduce
at least the high leve lconcepts and ideas behind our architecture as
a whole.

T-REX architecture is structured around the notion of 
{\em state variables} as the unique basis of interactionalong
with a well defined ownership  model for each of these. 
In this framework each reactor is seen as a
black box providing its own {\em internal} state variables for which
its is reponsible to maintain its value at every single instant --
represented by a {\em tick} with a fixed duration. In order to
identify its state a reactor can subscribe to {\em external} state
variables for  which it will received new updates when the owner of
these will produce its {\em internal} state. This {\em
  synchronization} process progate through the reactor dependency 
graph on a bottom up flow at every {\em tick} ensuring that all the
reactors have a consistent view of the world. Conversely each reactor
can request a future change on one of its {\em external} state
varaiable which will be transfromed as an {\em internal} goal to the
owner. This goal is expected to porduce a deliberation phase on the
receiver which can in turn produce a plan impacting its {\em external}
state variables that will propagate down following the same
mechanism.

@@

Overal T-REX architecture helps to abstract to the reactor interaction
to only state variables. Each reactor is practically ignoring where
its {\em external} state information is managed as long as this
information is available and properly maintained so they can properly 
manage their own {\em internal} state information which in turn may be
used by other reactors. Such formalism also constrain reactors
interaction to state information exchanges which are two folds:
\begin{itemize}
\item {\em Observation on the current state values}: This information
  propagates up in the reactor graph at the synchronization phase 
  and occur at the beginning of every time steps (called
  ticks). During this phase the agent provides to each reactor 
  updates on their {\em external} state variables for this tick and
  ask them to compute their {\em internal} state variables values for
  this same tick. By doing so we propagate a consistent view of that
  state of the world throughout all the reactors as time advance.
\item {\em Goals request on future state}: This information propagates
  down in the reactor graph. It reflects that, in order to satisfy own
  of its {\em internal} objectives, a reactor may have a plan that
  rely on future values of one of its {\em external} state. When such
  a plan is identified we ensure that this part of the plan is given
  as a goal to the owner of the given {\em external(s)} state variable(s). 
\end{itemize}

The choice of representation of state variables, observations and
goals is directly derieved to the timleines and tokens from europa. 
such representation is ich enough to allow the excahnge of state
information in a formal and yet flexible way and its transaltion in
planning frameworks that have an explicit representation of time 
(such as but not limited to europa) is often trivial.

The overall execution cycle of the agent is also abstracted as some
continuous planning/deliberation cycle between all the
reactors. Indeed , at the agent interface level, each reactor provided
only two abstract calls that conceptualize their execution:
\begin{itemize}
\item \texttt{synchronize} that takes the last {\em external}
  observations as argument and will return the {\em internal}
  observations for this reactor or a failure report.
\item \texttt{step} that accept new goals requested (if any) on the
  {\em internal} timeline of this reactor and will execute one step of
  deliberation. The returned value indicates if a complete plan has
  been found and in that case will also provide the part of this plan
  applied to {\em external} timelines of this reactor.
\end{itemize}
Both of these two deliberative functions have different scope and
temporal constraints. The \texttt{synchronize} call is considered as
atomic and will focus only on state inference from the current tick by
identifying the current value of the {\em internal} state variables
from the latest updates on the {\em external} variables. This process
will then propagate up in the reactor graph at any single tick to
ensure that all reactors are aware of the current state of the
world. \texttt{step} relates to a single decision step of the planning
process which is focused on the future evolution of the state
variables and it can take multiple steps spanning other multiple
ticks for a reactor to produce a plan that is both complete and
valid. when such plan is identified its {\em external} part is then
dispatched to the corresponding reactors which in turn can start their
deliberation \texttt{step}s in order to satisfy these new
objectives. As a consequence physical execution will occur as these
goals and planning propagate down in the reactor graph until they
eventually become goals for the lowest level reactor which -- instead
of deliberating -- may just transform this goal in a physical command
execution.

The way the agent interleave planning \texttt{step}s and
synchronization is illustrated in Figure \ref{fig:tick-exec}.
Synchronization happens at every single tick and will interrupt
planning allowing this reactor to identify its current state that can
propagate through the reactor graph in order to ensure t aconsitent
view of the state of the world between all the reactors.  
\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \includegraphics[width=0.55\columnwidth]{figs/tick-cycle}
  \caption{\small \texttt{T-REX} reactor execution cycle: {\em
      Deliberation} is interrupted by {\em synchronization} at the
    beginning of every {\em tick} allowing integration of state
    information.}
  \label{fig:tick-exec}
  \vskip-0.8pc
\end{figure}
This view of the reactor has two parallel tasks aloow for long
planning deliberation while ensuring that the state evolution
propagates through the agent as time advance. These two tasks 
intertwined together give the two natural flows (bottom-up for state
estimation and top to bottom for future plan) that we expect to see in
a control loop. As they are both conceptualized as inference processes
it is natural to implement a reactor based on generic automated
planning. For example in Fig. \ref{fig:agent}, appart \textsf{Shore
  comm.} and \textsf{Vehicle} acting as interfaces to the system,
every other reactors (colored in blue) are different instance of our
europa reactor differntiated only by their model and europa solver
configurations.

@@

The europa reactor is our implementation inside T-REX of a reactor
that usess automated planning for its execution. This reactor is fully
based on the europa framework and applies the core idea we of T-REX we
described above to implement a reactor for which the sole purpse is to
deliberate as time advance. The choice of europa as the planning
framework was directed by the fact thta this framework provide  a rich
and highly configurable framework to be able to infer around a rich
temporal domain.

This reactor is implemented following the idea that in T-REX, the
whole execution cycle is two deliberation processes. Both of these
processes are implemented as two europa solvers modifying the same
plan database and for which the execution is managed by the T-REX 
framework as time advance during the execution of the system:
\begin{enumerate}
\item The synchronization solver is a specialized europa solver that
  will integrate in the plan new information about the evolution of
  the External state variable and ensure that the reactor propagate
  these in order to both identify its current state and inform
  other reactors of any state change on its internal timelines. This
  solver is summoned at the beginning of every single tick.
\item The deliberation solver is managing the deliberation process of
  the reactor in order to produce a new plan or alter its current plan
  as new goals are given to the reactor or the synchronization solver
  identifying a conflict between the current state of the world and
  the expectations of the previous plan. This process can  span other
  multiple tick and therefore can be interrupted at any single tick in
  order for the synchronization solver to do its task.
\end{enumerate}
While the two processes are separate, they share the same plan
internal to a reactor and therefore at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when the planning process eventually finds a
solution, synchronization is informed about generating the planned
state values on the {\em external} state variables. This {\em
  external} plan defines the set of {\em goals} for this state
variable managed by another {\em reactor}.

An important challenge for embedding automayted planning i na situated
agent as always been to go around the asumptions made by most classical
planners. Specifically one assumption restrict the problem to ``offline
planning''  defined in \cite{ghallab04} as follow:
\begin{quotation}
  The planner is not concerned with any change that may occur in
  $\Sigma$ \footnote{In this book $\Sigma$ identifies the world modeled
    by the plan domain} while it is planning; it plans for the given
  initial and goal states regardless of the current dynamics, if any.
\end{quotation}
This assumption, while helpfull to reduce the scope of the planning
problem, becomes problematic if one wants to embed a planner in a
real-time system that evolves in a highly dynamic environment. In a
typical robot this implies that the only point where the planner can
be informed about the current world state is righe before the planning
is started by giving to the planner the intial state and current
objectives. During the planning, it is assumed that the agent can
maintin the world as perceived by the planner as stable. In 
\cite{lemai04, lemai-chenevier2004}, they do attempt to reduce the
impact of planning by aloowing local plan repairs which attempt to
insert repairing actions in the exisiting plan andf can be done fast
enough. Still in ythe eventuality that such repair is not possible
they add to stop their vehicle until the full (re-)planning was
complete. {\em I may need more examples/refs here}

@@

In our case though the interaction between planning and execution
does not stop at this level. A core point of the overal schedulling of
both {\em synchronization} and {\em planning} steps resides on our
architecture enforsincing that the synchroniuzation process can occur
in between any step of the plan. This lead us to the fact that not
only synchronization is imnfluenced by the outcome of planning but
this influence goes also in the alternate direction. Specifically when
synchronuiation occuers before we found a complete plan, it will
inject as we described new facts (the {\em external} observations and
resulting {\em internal} state) in the plan database. This
informatiuon will be present when the next planning \texttt{step} will
resume. Therefore they are not intergal part of the plan resolution
(generating potentially new flaws for the planner or contributing to
resolve previously exisiting flaws). The planning process is therefore
perturbed internallyand fully informed on how the world evolves as it
is planning.  This aspect trueluy contrast to other approaches we have
idenitifed in the literature which avoid to perturb the planner during
it search as it is not compatible with the ``off-line planning'' assumption.

By our design choices -- both at the architecture level and the integration of
the europa framework for embedded deliberation -- we were able to
implement an architecture that avoid this limitation while remaining in
a formal frame that still allow for deliberation and model-based
agent control. We believe that this tighter integration of planning
and execution allow the system to ba more accute of its environament
alowing better informaed decision and tighter reaction. This is
possible while still integrating information such as a long latency
for the reactors as long as one can ensure that synchronization of al
the reactors and deliberation steps can be done in a time that is
resonably small in comparison of the agent tick duration.

>>>>

The \rx architecture was designed with the objective to bring planning
closer to the low-level control of the vehicle while maintaining its
reactivity. Indeed, in most of the architecture where task planning is
embedded, this task remains at a failry high level of abstraction as
the complexity of this task would otherwise impede the system
reactiveness. As a result the planning problem managed {\em in situ}
remains fairly detached from low-level control and most of the
reactive adaptation which in contrast is managed by specialized
component that relies on different representations for modeling the
local behavior they control or even do not have any explicit formal
model of this behavior. The two angles our architecture attack this issue 
are the following:

\begin{itemize}

\item Distribute the global planning problem into multiple decision
  loops, called {\em reactors}. Each reactor have its own scope both 
  functionally (one for example focusing on the conversion of high 
  level waypoints into low-level commands) and temporally as 
  each reactor specifies to the agent their planning look-ahead and
  their expected latency before being able to produce plan.

\item Enforce a stronger interaction between the planning process 
  and the integration of the world evolution. The objective being to
  make sure that the plannign task is done while continuously informed 
  on new world state change. By doing so, we expect to allow the planner
  tohave better informed decision despite the fact that its latency is
  larger that the world change rate.

\end{itemize}

For the first point \rx provides a formal framework described in
details in \cite{py10} and \cite{rajan12} and is not the focus of this
paper. The second have deep implication on how on should integrate a
planner in this framework in general and was applied to the europa
framework specifically. We'll develop later how this requirement
implies to see in-situ planning as two parallels processes that are
tightly linked through the same plan structure. But before that we
need to introduce at least the high leve lconcepts and ideas behind
our architecture as a whole.

\rx architecture is structured around the notion of {\em state
  variables} as the unique basis of interactionalong with a well
defined ownership model for each of these.  In this framework each
reactor is seen as a black box providing its own {\em internal} state
variables for which its is reponsible to maintain its value at every
single instant -- represented by a {\em tick} with a fixed
duration. In order to identify its state a reactor can subscribe to
{\em external} state variables for which it will received new updates
when the owner of these will produce its {\em internal} state. This
{\em synchronization} process progate through the reactor dependency
graph on a bottom up flow at every {\em tick} ensuring that all the
reactors have a consistent view of the world. Conversely each reactor
can request a future change on one of its {\em external} state
variable which will be transfromed as an {\em internal} goal to the
owner. This goal is expected to porduce a deliberation phase on the
receiver which can in turn produce a plan impacting its {\em external}
state variables that will propagate down following the same mechanism.

@@

Overall the \rx architecture helps to abstract to the reactor
interaction to only state variables. Each reactor is practically
ignoring where its {\em external} state information is managed as long
as this information is available and properly maintained so they can
properly manage their own {\em internal} state information which in
turn may be used by other reactors. Such formalism also constrain
reactors interaction to state information exchanges which are two
folds:

\begin{itemize}

\item {\em Observation on the current state values}: This information
  propagates up in the reactor graph at the synchronization phase 
  and occur at the beginning of every time steps (called
  ticks). During this phase the agent provides to each reactor 
  updates on their {\em external} state variables for this tick and
  ask them to compute their {\em internal} state variables values for
  this same tick. By doing so we propagate a consistent view of that
  state of the world throughout all the reactors as time advance.

\item {\em Goals request on future state}: This information propagates
  down in the reactor graph. It reflects that, in order to satisfy own
  of its {\em internal} objectives, a reactor may have a plan that
  rely on future values of one of its {\em external} state. When such
  a plan is identified we ensure that this part of the plan is given
  as a goal to the owner of the given {\em external(s)} state variable(s).

\end{itemize}

The choice of representation of state variables, observations and
goals is directly derieved to the timelines and tokens from europa. 
such representation is rich enough to allow the exchange of state
information in a formal and yet flexible way and its transaltion in
planning frameworks that have an explicit representation of time (such
as but not limited to europa) is often trivial.

The overall execution cycle of the agent is also abstracted as a
continuous planning/deliberation cycle between all the
reactors. At the agent interface level, each reactor provided
only two abstract calls that conceptualize their execution:

\begin{itemize}

\item \texttt{synchronize} that takes the last {\em external}
  observations as argument and will return the {\em internal}
  observations for this reactor or a failure report.

\item \texttt{step} that accept new goals requested (if any) on the
  {\em internal} timeline of this reactor and will execute one step of
  deliberation. The returned value indicates if a complete plan has
  been found and in that case will also provide the part of this plan
  applied to {\em external} timelines of this reactor.

\end{itemize}

Both of these two deliberative functions have different scope and
temporal constraints. The \texttt{synchronize} call is considered as
atomic and will focus only on state inference from the current tick by
identifying the current value of the {\em internal} state variables
from the latest updates on the {\em external} variables. This process
will then propagate up in the reactor graph at any single tick to
ensure that all reactors are aware of the current state of the
world. \texttt{step} relates to a single decision step of the planning
process which is focused on the future evolution of the state
variables and it can take multiple steps spanning other multiple ticks
for a reactor to produce a plan that is both complete and valid. when
such plan is identified its {\em external} part is then dispatched to
the corresponding reactors which in turn can start their deliberation
\texttt{step}s in order to satisfy these new objectives. As a
consequence physical execution will occur as these goals and planning
propagate down in the reactor graph until they eventually become goals
for the lowest level reactor which -- instead of deliberating -- may
just transform this goal in a physical command execution.

The way the agent interleave planning \texttt{step}s and
synchronization for a single reactor is illustrated in Figure 
\ref{fig:tick-exec}. Synchronization happens at every single 
tick and will interrupt planning allowing this reactor to identify
 its current state that can propagate through the reactor graph
 in order to ensure t aconsitent view of the state of the world 
between all the reactors.

\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \includegraphics[width=0.55\columnwidth]{figs/tick-cycle}
  \caption{\small \texttt{T-REX} reactor execution cycle: {\em
      Deliberation} is interrupted by {\em synchronization} at the
    beginning of every {\em tick} allowing integration of state
    information.}
  \label{fig:tick-exec}
  \vskip-0.8pc
\end{figure}

This view of the reactor has two parallel tasks allow for long
planning deliberation while ensuring that the state evolution
propagates through the agent as time advance. These two tasks 
intertwined together give the two natural flows (bottom-up for state
estimation and top to bottom for future plan) that we expect to see in
a control loop. As they are both conceptualized as inference processes
it is natural to implement a reactor based on generic automated
planning. For example in Fig. \ref{fig:agent}, appart \textsf{Shore
  comm.} and \textsf{Vehicle} acting as interfaces to the system,
every other reactors (colored in blue) are different instance of our
europa reactor differntiated only by their model and europa solver
configurations.

@@

The \eu reactor is our implementation inside \rx of a reactor that
usess automated planning for its execution. This reactor is fully
based on the europa framework and applies the core idea of \rx we
described above to implement a reactor for which the sole purpose is to
deliberate as time advance. The design of this reactor takes cues from 
the IDEA architecture presented on \cite{mus02, mus06} in the sense taht both
planning and execution are based on a single unified model. The choice 
of \eu as the planning framework was directed by the fact that this 
framework provide a rich and highly configurable toolset to be able to 
infer around a rich temporal domain.

This reactor is implemented following the idea that in \rx, the
whole execution cycle is two deliberation processes. Both of these
processes are implemented as two europa solvers modifying the same
plan database and for which the execution is managed by the \rx 
framework as time advance during the execution of the system:

\begin{enumerate}

\item The synchronization solver is a specialized europa solver that
  will integrate in the plan new information about the evolution of
  the External state variable and ensure that the reactor propagate
  these in order to both identify its current state and inform
  other reactors of any state change on its internal timelines. This
  solver is summoned at the beginning of every single tick.

\item The deliberation solver is managing the deliberation process of
  the reactor in order to produce a new plan or alter its current plan
  as new goals are given to the reactor or the synchronization solver
  identifying a conflict between the current state of the world and
  the expectations of the previous plan. This process can  span other
  multiple tick and therefore can be interrupted at any single tick in
  order for the synchronization solver to do its task.

\end{enumerate}

While the two processes are separate, they share the same plan
internal to a reactor and therefore at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when the planning process eventually finds a
solution, synchronization is informed about generating the planned
state values on the {\em external} state variables. This {\em
  external} plan defines the set of {\em goals} for this state
variable managed by another {\em reactor}.

An important challenge for embedding automayted planning i na situated
agent as always been to go around the asumptions made by most classical
planners. Specifically one assumption restrict the problem to ``offline
planning''  defined in \cite{ghallab04} as follow:

\begin{quotation}
  The planner is not concerned with any change that may occur in
  $\Sigma$ \footnote{In this book $\Sigma$ identifies the world modeled
    by the plan domain} while it is planning; it plans for the given
  initial and goal states regardless of the current dynamics, if any.
\end{quotation}

This assumption, while helpfull to reduce the scope of the planning
problem, becomes problematic if one wants to embed a planner in a
real-time system that evolves in a highly dynamic environment. In a
typical robot this implies that the only point where the planner can
be informed about the current world state is right before the planning
is started by giving to the planner the intial state and current
objectives. During the planning, it is assumed that the agent can
maintin the world as perceived by the planner as stable. In
\cite{lemai04, lemai-chenevier2004}, they do attempt to reduce the
impact of planning by allowing local plan repairs which attempt to
insert repairing actions in the exisiting plan andf can be done fast
enough. Still in ythe eventuality that such repair is not possible
they add to stop their vehicle until the full (re-)planning was
complete. {\em I may need more examples/refs here}

@@

In our case though the interaction between planning and execution does
not stop at this level. A core point of the overal schedulling of both
{\em synchronization} and {\em planning} steps resides on our
architecture enforsincing that the synchroniuzation process can occur
in between any step of the plan. This lead us to the fact that not
only synchronization is imnfluenced by the outcome of planning but
this influence goes also in the alternate direction. Specifically when
synchronuiation occuers before we found a complete plan, it will
inject as we described new facts (the {\em external} observations and
resulting {\em internal} state) in the plan database. This
informatiuon will be present when the next planning \texttt{step} will
resume. Therefore they are not intergal part of the plan resolution
(generating potentially new flaws for the planner or contributing to
resolve previously exisiting flaws). The planning process is therefore
perturbed internallyand fully informed on how the world evolves as it
is planning.  This aspect trueluy contrast to other approaches we have
identified in the literature which avoid to perturb the planner during
it search as it is not compatible with the ``off-line planning''
assumption.

By our design choices -- both at the architecture level and the
integration of the europa framework for embedded deliberation -- we
were able to implement an architecture that avoid this limitation
while remaining in a formal frame that still allow for deliberation
and model-based agent control. We believe that this tighter
integration of planning and execution allow the system to ba more
acute of its environment allowing better informaed decision and
tighter reaction. This is possible while still integrating information
such as a long latency for the reactors as long as one can ensure that
synchronization of al the reactors and deliberation steps can be done
in a time that is resonably small in comparison of the agent tick
duration.
